Hypertension and microbiome

### 2019-05-24 articletwo.Rmd
Missing commas
PCoA analysis and axis 1+3

@@

@@

```{r pcoa plot}
pcoa.vectors <- pcoa.ordinate$vectors %>% as.data.frame
eig.fracs <- pcoa.ordinate$values$Relative_eig
pcoa.df <- merge(pcoa.vectors %>% select(Axis.1, Axis.2),
                 meta(pseq.species), by=0, all = TRUE) %>%
    mutate(STATETREATMENT = factor(paste0(HYPERTENSION, ANYDRUG),
                                   levels = c("11", "10", "01", "00")))

pcoa.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y=Axis.2, color = STATETREATMENT, shape = STATETREATMENT)) +
    geom_point(size=0.4, alpha=0.4) +
    xlab(sprintf("PCoA Axis 1 (%.1f%%)", 100*eig.fracs[1])) +
    ylab(sprintf("PCoA Axis 2 (%.1f%%)", 100*eig.fracs[2])) +
    scale_colour_manual(name = "State and treatment",
                        labels = c("Hypertensive, medication",
                                   "Hypertensive, no medication",
                                   "Normotensive, no medication"),
                        breaks=c("11", "10", "00"),
                        values = c("red", "red", "blue", "blue")) +   
    scale_shape_manual(name = "State and treatment",
                       labels = c("Hypertensive, medication",
                                  "Hypertensive, no medication",
                                  "Normotensive, no medication"),
                       breaks = c("11", "10", "00"),
                       values = c(4, 1, 1))+
    coord_fixed() +
    scale_x_continuous(breaks=seq(-100, 100, 20)) +
    scale_y_continuous(breaks=seq(-30, 30, 10)) +
    theme_classic() +
    guides(colour = guide_legend(override.aes = list(size=4))) +
    theme(text = element_text(size=14),
          legend.title = element_blank(),
          legend.text=element_text(size=10),
          legend.justification=c(1,0),
          legend.position=c(1,0),
          legend.margin=margin(t = 1, unit='mm'),
          legend.background = element_rect(fill=alpha('gray', 0.1)))

ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=5, width=5, dpi = 300)
```

```{r Write figure 2, include=FALSE}
fig2head <- "Principal coordinate analysis for bacterial abundances."
fig2foot <- sprintf("Principal coordinate analysis for bacterial abundances at species level. Bray-Curtis distances are calculated for abundances after compositional transformation. First two axes explain %.1f%% of the variation.", round(100*sum(eig.fracs[1:2]),1))
writeimage(doc, 2, "cache/pcoa-species.png", fig2head, fig2foot)
```

@@

@@

```{r odds htn lacto}
odds.hypertension.lactogroup <- glm.hypertension.lactogroup %>%
    filter(!grepl("Intercept", term)) %>%
    select(term, estimate, conf.low, conf.high) %>%
    add_row(term = "lacto_groups1", estimate = 1.0, conf.low = NA, conf.high = NA) %>%
    dplyr::filter(grepl("lacto", term))
odds.hypertension.lactogroup %>% kable
```

>>>>

## Firmicutes bacteroides ratio

Firmicutes:Bacteroidetes ratio

```{r firmicutes bacteroides, eval = FALSE}
bfratio <- function (x) {
    a <- transform(aggregate_taxa(x, level = "Phylum"), "compositional")
    b <- abundances(a)["Bacteroidetes",]
    f <- abundances(a)["Firmicutes",]  
    #data.frame(Bacteroidetes = b, Firmicutes = f, Ratio = b/f)
    b/f
}

packageVersion("microbiome")

bfratio(pseq.species)

phy3 <- aggregate_taxa(pseq.species, level = "Phylum")
aggregate_taxa

 transform(aggregate_taxa(x, level = "Phylum"), "compositional")
sessionInfo()

@@

Results in minimum model of alpha diversity

```{r talbe alpha div min}
alphadiversity$min %>%
    select(response, estimate, p.value, conf.low, conf.high) %>%
    mutate(mean_ci = sprintf("p=%s; 95%%-CI %.2f-%.2f",
                              pub.p(p.value), conf.low, conf.high)) %>%
    mutate_if(is.numeric, round, 3) %>%
    arrange(p.value) %>%
    kable
```

Results for full model of alpha diversity

```{r talbe alpha div max}
alphadiversity$max %>%
    select(response, estimate, p.value, conf.low, conf.high) %>%
    mutate(mean_ci = sprintf("p=%s; 95%%-CI %.2f-%.2f",
                              pub.p(p.value), conf.low, conf.high)) %>%
    mutate_if(is.numeric, round, 3) %>%
    arrange(p.value) %>%
    kable
```

Results for beta diversity

```{r table beta diversity}
adonis.species$max %>%
    data.table::rbindlist(id = "model_name") %>%
    dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) %>%
    dplyr::select(response, R2, "Pr(>F)") %>%
    mutate(p.value = pub.p(`Pr(>F)`))
```

@@

```{r pcoa plot}
pcoa.vectors <- pcoa.ordinate$vectors %>% as.data.frame
eig.fracs <- pcoa.ordinate$values$Relative_eig
pcoa.df <- merge(pcoa.vectors %>% select(Axis.1, Axis.2, Axis.3),
                 meta(pseq.species), by=0, all = TRUE) %>%
    mutate(STATETREATMENT = factor(paste0(HYPERTENSION, ANYDRUG),
                                   levels = c("11", "10", "01", "00")),
           gMAP = cut(oMAP, seq(60, 160, 15)))

pcoa.df %>% pull(HYPERTENSION) %>% summary

pcoa.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y=Axis.3, color = HYPERTENSION)) +
    geom_point(size=0.1, alpha=0.4) +
    xlab(sprintf("PCoA Axis 1 (%.1f%%)", 100*eig.fracs[1])) +
    ylab(sprintf("PCoA Axis 3 (%.1f%%)", 100*eig.fracs[3])) +
    scale_colour_manual(name = "State and treatment",
                        labels = c("Normotensive", "Hypertensive"),
                        breaks=c(0, 1),
                        values = c("blue", "red")) +   
    coord_fixed() +
    scale_x_continuous(breaks=seq(-1, 1, .25)) +
    scale_y_continuous(breaks=seq(-1, 1, .25)) +
    theme_classic() +
    guides(colour = guide_legend(override.aes = list(size=4))) +
    theme(text = element_text(size=14),
          legend.title = element_blank(),
          legend.text=element_text(size=10),
          legend.justification=c(1,0),
          legend.position=c(1,0),
          legend.margin=margin(t = 1, unit='mm'),
          legend.background = element_rect(fill=alpha('gray', 0.1)))

ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=4, width=5, units = "in", dpi = 300)
```

First three axis explain portion (`r eig.fracs[1:3]`) of
variation. Testing associations to blood pressure variables on PCoA
axis 1

```{r pcoa axis analysis axis 1}
calculateglm(pcoa.df,
             modelstring = "%s ~ Axis.1 + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
             filterstr = "Axis") %>% 
    select(response, term, estimate, std.error, p.value) %>%
    mutate_if(is.numeric, round, 3) %>%
    kable
```

Testing associations to blood pressure variables on PCoA axis 2

```{r pcoa axis analysis axis 2}
calculateglm(pcoa.df,
             modelstring = "%s ~ Axis.2 + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
             filterstr = "Axis") %>% 
    select(response, term, estimate, std.error, p.value) %>%
    mutate_if(is.numeric, round, 3) %>%
    kable
```

Testing associations to blood pressure variables on PCoA axis 3

```{r pcoa axis analysis axis 3}
calculateglm(pcoa.df,
             modelstring = "%s ~ Axis.3 + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
             filterstr = "Axis") %>%
    select(response, term, estimate, std.error, p.value) %>%
    mutate_if(is.numeric, round, 3) %>%
    kable
```

```{r Write figure 2, include=FALSE}
fig2head <- "Principal coordinate analysis for bacterial abundances."
fig2foot <- sprintf("Principal coordinate analysis for bacterial abundances at species level. Bray-Curtis distances are calculated for abundances after compositional transformation. First two axes explain %.1f%% of the variation.", round(100*sum(eig.fracs[c(1,3)]),1))
writeimage(doc, 2, "cache/pcoa-species.png", fig2head, fig2foot)
```

@@

```{r maaslin table}
maas.results.df.signf %>%
    select(Feature, Value, Coefficient, Qval) %>%
    arrange(Qval) %>%
    kable
```

@@

```{r odds htn lacto}
odds.hypertension.lactogroup <- glm.hypertension.lactogroup %>%
    filter(!grepl("Intercept", term)) %>%
    select(term, estimate, p.value, conf.low, conf.high) %>%
    add_row(term = "lacto_groups1", estimate = 1.0, conf.low = NA, conf.high = NA) %>%
    dplyr::filter(grepl("lacto", term))
odds.hypertension.lactogroup %>% kable
```
